An Analysis of Security Vulnerabilities in Container Images for Scientific Data Analysis

### 2020-10-26 paper.tex

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/results.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected by
Anchore and Stools in container images. \textbf{(A)} Number of vulnerabilities by
container image and severity, showing hundreds of detected vulnerabilities
per image. Images s*,t*,u*,v*,w* and x* are Singularity images scanned by Stools. \textbf{(B)} Effect of image minification and
package update on 5 container images, showing that both techniques are
complementary \textbf{(C)} Number of vulnerabilities by number of
packages, showing a strong linear relationship \tristan{add Singularity images and update regression parameters accordingly.}. \textbf{(D)} Number of
vulnerabilities by number of packages \emph{after package update}, showing that software updates 
importantly reduce the number of detected vulnerabilities.}
\end{figure*}

An important amount of vulnerabilities were found in the tested container
 images (Fig~\ref{fig:vulnerabilities}-\textbf{A}), with an average of 468 vulnerabilities
  per image and a median of 372 \tristan{update values with singularity images}. In comparison, no vulnerabilities were found in base
 Docker images \texttt{ubuntu:20.04} and \texttt{centos:7} after package
 update. Moreover, a significant fraction of detected vulnerabilities are
 of high severity
 (\href{https://www.first.org/cvss/specification-document}{CVSS} score
 >=7.0) and a few of them are of critical severity (CVSS >= 9.0). Remote
 attackers could possibly exploit these vulnerabilities to execute
 arbitrary code in the container, by crafting responses to specific network
 requests. Images based on the Alpine distribution 
 had the lowest numbers of vulnerabilities, but no significant difference
 in the numbers of vulnerabilities detected in 
 Ubuntu, Debian or CentOS distributions was observed.

Unsurprisingly, a strong linear relationship is found between the number of
detected vulnerabilities and the number of packages present in the
image (Fig~\ref{fig:vulnerabilities}-\textbf{C}, r=0.83,
$p\textless10^{-10}$) \tristan{update values with singularity images}. On average, 1.8 vulnerabilities are introduced for
each new package installation. This observation motivates a systematic
review of software dependencies by application developers, to avoid
unnecessary packages in container images. This is also an argument in favor of lightweight
distributions such as Alpine. Compared to Ubuntu and Debian distributions,
CentOS images seem to have a lower number of vulnerabilities by package on
average, although data is too scarce to conclude.

@@

Updating container images reduces the number of vulnerabilities by package
by a factor of 3 on average, resulting in only 0.6 extra vulnerabilities by
package (Fig~\ref{fig:vulnerabilities}-\textbf{D}, r=0.81,
$p\textless10^{-7}$). Twelve container images are missing on this figure:
six of them could not be updated due to various issues with the package
manager, and six of them are Singularity images that we didn't update.
% \change{3 out of these missing 6 images retured an error while updating which
% indicates a problem with the package installer, 2 were unable to update because they
% reached EOL and last one with CentOS 7.1.1503 returned libselinux conflicts with systemd} 
In spite of the associated reproducibility challenges, updating
packages therefore appears to be an efficient way to avoid vulnerabilities. It
is not an ultimate solution though, as a substantial number of
vulnerabilities remain.

>>>>

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/results.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected by
Anchore and Stools in container images. \textbf{(A)} Number of vulnerabilities by
container image and severity, showing hundreds of detected vulnerabilities
per image. Images s*,t*,u*,v*,w* and x* are Singularity images scanned by Stools and
others are Docker images scanned using Anchore. \textbf{(B)} Effect of image minification and
package update on 5 container images, showing that both techniques are
complementary \textbf{(C)} Number of vulnerabilities by number of
packages, showing a strong linear relationship. \textbf{(D)} Number of
vulnerabilities by number of packages \emph{after package update}, showing that software updates 
importantly reduce the number of detected vulnerabilities.}
\end{figure*}

An important amount of vulnerabilities were found in the tested container
 images (Fig~\ref{fig:vulnerabilities}-\textbf{A}), with an average of 460 vulnerabilities
  per image and a median of 321. In comparison, no vulnerabilities were found in base
 Docker images \texttt{ubuntu:20.04} and \texttt{centos:7} after package
 update. Moreover, a significant fraction of detected vulnerabilities are
 of high severity
 (\href{https://www.first.org/cvss/specification-document}{CVSS} score
 >=7.0) and a few of them are of critical severity (CVSS >= 9.0). Remote
 attackers could possibly exploit these vulnerabilities to execute
 arbitrary code in the container, by crafting responses to specific network
 requests. Images based on the Alpine distribution 
 had the lowest numbers of vulnerabilities, but no significant difference
 in the numbers of vulnerabilities detected in 
 Ubuntu, Debian or CentOS distributions was observed.

Unsurprisingly, a strong linear relationship is found between the number of
detected vulnerabilities and the number of packages present in the
image (Fig~\ref{fig:vulnerabilities}-\textbf{C}, r=0.82,
$p\textless10^{-12}$). On average, 1.7 vulnerabilities are introduced for
each new package installation. This observation motivates a systematic
review of software dependencies by application developers, to avoid
unnecessary packages in container images. This is also an argument in favor of lightweight
distributions such as Alpine. Compared to Ubuntu and Debian distributions,
CentOS images seem to have a lower number of vulnerabilities by package on
average, although data is too scarce to conclude.

@@

Updating container images reduces the number of vulnerabilities by package
by a factor of 3 on average, resulting in only 0.6 extra vulnerabilities by
package (Fig~\ref{fig:vulnerabilities}-\textbf{D}, r=0.81,
$p\textless10^{-8}$). Twelve container images are missing on this figure:
six of them could not be updated due to various issues with the package
manager, and six of them are Singularity images that we didn't update.
% \change{3 out of these missing 6 images retured an error while updating which
% indicates a problem with the package installer, 2 were unable to update because they
% reached EOL and last one with CentOS 7.1.1503 returned libselinux conflicts with systemd} 
In spite of the associated reproducibility challenges, updating
packages therefore appears to be an efficient way to avoid vulnerabilities. It
is not an ultimate solution though, as a substantial number of
vulnerabilities remain.
